Adaptative Computation:\\ An Essay in Memory of John H. Holland

### 2016-02-09 paper.tex
made axelrod edits, except for adding more refs to markets section

The SFI Artificial Stock Market project \cite{Arthur1997b,Palmer1994}
developed an exploratory model in which rational
traders were replaced by \emph{adaptive} traders---those who learn to
forecast stock prices over time.  The model tested for the
emergence of different trading strategies, including
\emph{fundamental}, \emph{technical}, or \emph{uninformed}
strategies. The simulated market with adaptive trading agents was run
many times, and the dynamics of price and trading volumes were
compared to observed patterns in real markets.  Holland and his
collaborators found that the model's dynamics replicated several
features of real-life markets.

@@

Holland's focus on understanding the mechanisms by which complex patterns emerge and change,
rather than simply characterizing the patterns themselves (e.g.,
describing chaotic attractors or power laws), reflected his
determination to get to the heart of complex adaptive systems.
This determination represents the best of science.  Holland's
willingness to tackle the hardest questions, develop his own
formalisms, and use mathematics to provide insight sets a high bar for
scientists in all disciplines.
%to whch we
%all should aspire.

@@

Behind the ideas, of course, is the man himself.   Everyone who knew
John personally will remember the gleam in his eye when he
encountered a new idea; his willingness to talk to anyone, no matter
how famous or obscure; and his incredible generosity and patience.
%His insights aboutwere deeper and more general than what often passes for work in biomimicry, e.g., for robots.
Holland's personality and humanity were somehow inextricably tangled
up with his intellectual contributions.  Since his death, many of
Holland's former students and colleagues have movingly described their
desire to emulate his personal qualities as much as his scientific
excellence.  John Holland's ideas, intellectual passion, and personal approach will serve as
beacons for research in intelligent and complex systems for many
years to come. 
%the world of science is poorer without John
%Holland on the scene.

@@

The authors thank L. Booker, R. Riolo, K. Winter, and the anonymous
reviewers for their careful reading of the manuscript and many helpful
suggestions.  SF gratefully acknowledges the partial support of NSF
(NeTS 1518878, CNS 1444500), DARPA (FA8750-15-C-0118), 
Air Force Office of Scientific Research (FA8750-15-2-0075), and the Santa Fe Institute.

>>>>

The SFI Artificial Stock Market project \cite{Arthur1997b,Palmer1994}
developed an exploratory model in which rational
traders were replaced by \emph{adaptive} traders---those who learn to
forecast stock prices over time.  The model tested for the
emergence of different trading strategies, including
\emph{fundamental}, \emph{technical}, or \emph{uninformed}
strategies. The simulated market with adaptive trading agents was run
many times, and the dynamics of price and trading volumes were
compared to observed patterns in real markets.  Holland and his
collaborators found that the model's dynamics replicated several
otherwise puzzling features of real-life markets.

@@

Holland's focus on understanding the mechanisms by which complex patterns emerge and change,
rather than simply characterizing the patterns themselves (e.g.,
describing chaotic attractors or power laws), reflected his
determination to get to the heart of complex adaptive systems.
This determination represents the best of science.  Holland's
willingness to tackle the hardest questions, develop his own
formalisms, and use mathematics and simulation to provide insight sets a high bar for
scientists in all disciplines.
%to whch we
%all should aspire.

@@

Behind the ideas, of course, is the man himself.   Everyone who knew
John personally will remember the gleam in his eye when he
encountered a new idea; his willingness to talk to anyone, no matter
how famous or obscure; and his incredible generosity and patience.
%His insights aboutwere deeper and more general than what often passes for work in biomimicry, e.g., for robots.
Holland's personality and humanity were somehow inextricably entangled
with his intellectual contributions.  Since his death, many of
Holland's former students and colleagues have movingly described their
desire to emulate his personal qualities as much as his scientific
excellence.  John Holland's ideas, intellectual passion, and personal approach will serve as
beacons for research in intelligent and complex systems for many
years to come. 
%the world of science is poorer without John
%Holland on the scene.

@@

The authors thank R. Axelrod, L. Booker, R. Riolo, K. Winter, and the anonymous
reviewers for their careful reading of the manuscript and many helpful
suggestions.  SF gratefully acknowledges the partial support of NSF
(NeTS 1518878, CNS 1444500), DARPA (FA8750-15-C-0118), 
Air Force Office of Scientific Research (FA8750-15-2-0075), and the Santa Fe Institute.
